---
title: "Assignment 3"
author: "Arthur Junges Schmidt"
date: "`r as.character(Sys.time(), '%d %B %Y')`"
documentclass: article
lang: eng
classoption: a4paper
output:
  pdf_document:
    fig_width: 10
    fig_height: 6
geometry: margin=2cm


---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(e1071)
```

# Problem 1
(a) The file ps3-1.csv contains a data set with 34 features (x1, x2, … x34) and 1 target variable
(Y). Estimate a classifier model using Support Vector Machine and Random Forest algorithm in
R, respectively, with the first 14628 rows of the data. Optimize your model so that a False
Positive Rate is less than 10% for Y = 0 (actual Y = 1 cases falsely classified as Y = 0).
Particularly, you are required to review relevant literature, use the k-fold cross-validation method
to train the Random Forest model. Use grid search to find hyper-parameter setting: the best
number of trees and features, maximum leaf nodes. Assess importance of each feature based on
two criteria: Mean Decrease Accuracy and Mean Decrease Gini.
Compare two estimation methods with confusion matrices

```{r Data reading, message=FALSE}
Crude_Data <- read.csv("ps3-1.csv");
Crude_Data <- Crude_Data[,-1];
Crude_Data$Y <- factor(Crude_Data$Y, levels = c(0,1));
Training_Data <- Crude_Data[1:14628, ];
Test_Data <- Crude_Data[-(1:14628), ];
```


```{r SVM, cache=TRUE}
SVM_Model <- svm(Y ~ ., data = Training_Data)
#plot(SVM_Model, data = Training_Data, Y~.)
SVM_Predict <- predict(object = SVM_Model, newdata = Test_Data[,-35]);
table(Predicted = Test_Data[,35], True = SVM_Predict)

```

